Information Aggregation in Exponential Family Markets

### 2014-04-21 WriteUp/response.txt

- Reduce complications in notation: Please only keep necessary notation in this paper. @sebastien
  - For example, I found the reference to Bergman divergence a little annoying. I also think that maximum entropy principle is overemphasized---proving that the proposed scoring rule is proper is enough to me. I also got lost at the last paragraph of Section 1.

- intro revision: we meant this comment for the narrow case of MSR-style markets. @jake
  - Ostrovsky and of Wolfers and Zitzewitz are exactly about these issues. Pennock, in his Ph.D. thesis (1999), asks the question of what happens in equilibrium when traders have different priors and budget constraints, and shows that in equilibrium, the price can be interpreted as a budget-weighted average of trader beliefs. Beyond this, most of the empirical work on prediction markets is asking exactly the question of what prediction market prices actually mean. The recent working paper of Rothschild and Sethi ("Trading Strategies and Market Microstructure: Evidence from a Prediction Market") discusses how to interpret prices in the context of a continuous market rather than equilibrium prices -- the references in that paper are also worth looking at. Much of the work on manipulation is also implicitly focused on exactly this question (including the paper "Gaming Prediction Markets" with one of the authors of this paper as a co-author!)

- [COMPLETE] change rahul sami's affiliation
- [COMPLETE] remove parskip
- [COMPLETE] P4: please clarify the setup. Are u always using a single outcome x? same in the example section at Page 7 @sindhu 
  "changed sentence to: We consider a measurable space consisting of a set of (mutually exclusive, exhaustive) outcomes $\mX$ together with a $\sigma$-algebra $\mF$. i feel this change takes care of both these points"
- [COMPLETE] Please give a formal definition of exponential family distributions (with citation) @sindhu 
  "we didn't have a reference for this; i've added wainwright jordan."
- [COMPLETE] P7: U are using mu for both the report and expectation. Please use a different notation (e.g. hat(mu) @sindhu 
  "i think the reviewer means that in the definition of the scoring rule, we should indicate that the first argument is the agent report and not the true mean. so i've changed that."
- [COMPLETE] Section 6: 
  "Changed first paragraph of this section and added a sentence to the "Budget-limited trades" paragraph"
  - [COMPLETE] What exactly does the statement "the traders trade in multiple instances of the market" mean? 
  - [COMPLETE] The semantics of the time-varying budget alpha_i^t are not clear. @sindhu
  - [COMPLETE] What are the formal definitions of an "uninformative" and an "informative" trader? 
- [COMPLETE] I found the "underlying density" and "full density" discussion on P4 very confusing (what are these two densities?). @sindhu
  "I think he means p.6 just before thm 2.2; i've reworded this sentence."
- [COMPLETE] Regarding budgets and play money: indeed the play-money markets are a case where budget constraints are key. We have experience with this in other work where a play-money market was released for the US Elections. Here the budget constraints help alleviate the problem that traders can easily manipulate prices with unreasonable trades -- the goal of Section 6 was to give formal results to this effect. Risk aversion is also plausible with play-money; for instance, a trader high on the leaderboard may not want to risk her standing with large late trades. @sindhu
  "added a note to the introduction."
  - the example for risk-aversion of someone not wanting to lose their place on the leaderboard doesn't really hold water: that's a type of tournament incentive: very different from traditional risk-aversion.
- [COMPLETE] P6, THM 2.2. Where the sufficient statistics condition fits in. Can it be that the entropy maximizing exponential distribution is not like that. Please clarify. @sindhu
  "changed statement of thm 2.2 to state the relationship between \mu and \phi explicitly"
- [COMPLETE] examples of commonly studied distributions that can be modeled by the exponential family studied in this paper. @sindhu 
  " i added examples before prop 2.1 (normal, binomial and beta) and added a detailed mapping for the normal distribution"

@@

- P3: Section 2, after first paragraph. Please clarify that P contains ALL distributions on the outcome space. @sindhu
  "already says that?"
- P4, last sentence. Please quote what exactly was proven by [Grunwald and Dawid] @sindhu
  "i'm not sure about this one"
[COMPLETE]- The importance of the underlying measure v is understated. If v is always the Lebesgue measure then many distributions cannot be modeled by the exponential family as presented in the paper, e.g. normal distributions with fixed variance. Many discussions and proofs are made without reference to v. My question is: is it true that all the results and comments can be applied to arbitrary v? (I just checked a few but not all since many calculations are already oversimplified.) @sebastien: clarified that base measure is implicit in integrals.
[COMPLETE] -  Please give an example of non-Lebesgue v: explained the case of fixed variance, unknown mean in a new footnote.
[COMPLETE]- Can you model classical prediction markets with finite outcome space under the proposed framework? If so then what distribution in the exponential family its correspond to? @sebastien: added examples of scoring rules and prediction markets with finite outcome space, i.e., the usual log score and LMSR.
[COMPLETE] - Highlight key contributions: the authors could have just focused on presenting characterization of the class of new proper scoring rules and behavior and analysis of agents with exponential utility function. The connection between exponential family and the proposed prediction market setting is interesting and helpful if it is briefly mentioned and discussed in the end of the paper. The paper in its current form reads like suggesting that using exponential families is justified at the modeling level, which sounds a little weak (and in fact not necessary). So I dont think they should be mentioned in the beginning. For example, I got confused by "outstanding shares". Also not all exponential family distributions have a mean parameter (which is fine if you choose to model the market price distribution in this way, but this may limit the generality of the framework). @sebastien: added sentence explaining that we aren't actually modeling agent beliefs as exponential families in the scoring rule section.
[COMPLETE] - Section 5: Please clarify the setup up-front. What is known to the principal on the traders? Does it know $a$ or it can be estimated @sebastien: added sentence in Section 5 explaining market maker does not need to know a for results to hold.
[COMPLETE] - If accepted, the authors should spend some effort in discussing the connection of this work to that of Abernethy-Frongilllo (COLT 2012) 
- That paper shows how to construct a proper scoring rule for linear properties from any convex function using Bregman divergence. This paper seems to basically present a way of choosing that convex function; this is not uninteresting in itself, but I'm concerned as to why this connection is not made clearer in the paper. @sebastien: added discussion of all this right after Thm 2.2, and also used the right result from Aberneth-Frongillo'12 in the proof of Thm 2.2.

### 2014-04-21 WriteUp/maxentpredmkts.tex
addressed my reviewer comments

\begin{document}
% Page heads
\markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
% Title portion
\title{Information Aggregation in Exponential Family Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor}\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}
%%\author{GANG ZHOU
%%\affil{College of William and Mary}
%%YAFENG WU
%%\affil{University of Virginia}
%%TING YAN
%%\affil{Eaton Innovation Center}
%%TIAN HE
%%\affil{University of Minnesota}
%%CHENGDU HUANG
%%\affil{Google}
%%JOHN A. STANKOVIC
%%\affil{University of Virginia}
%%TAREK F. ABDELZAHER
%%\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
We consider the design of prediction market mechanisms known as automated market makers. We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a popular and well-studied probability distribution template used in statistics. We give a full development of this relationship and explore a range of benefits. We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that rely on exponential family distributions. We develop a very natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility. We also consider similar aspects under alternative models, such as when traders are budget constrained.
%
% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

@@

\terms{Algorithms, Economics}

\keywords{logarithmic score, exponential family, maximum entropy, risk aversion, budget constraints}

@@

Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event. Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event. Traders are allowed to report their beliefs by buying and selling securities whose ultimate payoff depends on the future outcome. This will affect the state of the market, thus updating the predictive probabilities for the event. Further, since the trades are done sequentially, the trader is allowed to observe all past trades in the market and update his private information based on this information. In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.

@@

In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations. Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{bayesian inference}---relate to natural aspects of our class of mechanisms. In particular, the central objects in our market framework can be interpreted via concepts used to define exponential families:
\begin{itemize}
	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter vector} of the distribution;
	\item the market prices correspond to \emph{mean parameters};
	\item the market's cost function corresponds to the distribution's \emph{log-partition} function.
%	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
\end{itemize}
% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}

We begin in Section~\ref{sec:scoring} with a discussion of scoring rules based on exponential family distributions, and we show how the framework leads to a variety of scoring rules for continuous outcome spaces.
We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms. In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well. In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
\begin{itemize}
	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists. In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
	\item In Section~\ref{sec:exputil} we consider \emph{risk-averse} agents that optimize their bets according to exponential utility. 
	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; this result is achieved via a potential game argument. The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are proportional to risk aversion parameters. 
	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market. We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
\end{itemize}

@@

Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces. Such rules have been characterized by~\citet{Savage71}; see also~\citep{frongillo2013elicitingmeans,lambert2008eliciting}. Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated. Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
% In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.

@@

We are interested in eliciting information about the agent's belief, in particular expectation information. Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite. The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome. A \emph{scoring rule} is a device for this purpose. Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.  A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$. The following definition is due to~\citet{lambert2008eliciting}.
%
\begin{definition} \label{def:proper-score}
A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have for all $\hmu \neq \mu$
%
\begin{equation} \label{eq-proper}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)].
\end{equation}
%
% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
\end{definition}
%

@@

It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions (see, for instance, \cite{WainJordan08}). These are distributions whose densities with respect to $\nu$ take the form
%
\begin{equation} \label{exp-fam}
p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
\end{equation}
%
The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with the linear constraints in~(\ref{maxent-prog}). The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
%
\begin{equation} \label{log-part}
T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
\end{equation}
%
The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family. Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space. The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular. The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) note
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. Note that the expected log score does not depend  on the entire maximum entropy distribution but just its expected statistic $\mu$. This is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$, where $\mu=\E_{p(x;\mu)}[\phi(x)]$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{gneiting2007strictly} and~\citet{Savage71}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%

As further intuition for the result, note that~(\ref{breg-div}) is the definition of the `Bregman divergence' with respect to strictly convex function $T$, written $D_T$. Therefore we have 
%
\begin{equation*}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] = D_T(\htheta,\theta) = D_G(\mu, \hmu),
\end{equation*}
%
where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$. The equation states that the agent's regret from misreporting its mean parameter does not depend on the full density $p$, only the mean $\mu$.

@@

Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations. Define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$. The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution. The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the usual log score.

>>>>

\begin{document}
% Page heads
\markboth{Abernethy et al.}{Information Aggregation in Exponential Family Markets}
% Title portion
\title{Information Aggregation in Exponential Family Markets}
\author{JACOB ABERNETHY
\affil{University of Michigan, Ann Arbor}
SINDHU KUTTY
\affil{University of Michigan, Ann Arbor}
S\'{E}BASTIEN LAHAIE
\affil{Microsoft Research, New York City}
RAHUL SAMI
\affil{University of Michigan, Ann Arbor\footnote{Current affiliation: Google, Inc. This work was done while the author was at the University of Michigan.}}}
%%\author{GANG ZHOU
%%\affil{College of William and Mary}
%%YAFENG WU
%%\affil{University of Virginia}
%%TING YAN
%%\affil{Eaton Innovation Center}
%%TIAN HE
%%\affil{University of Minnesota}
%%CHENGDU HUANG
%%\affil{Google}
%%JOHN A. STANKOVIC
%%\affil{University of Virginia}
%%TAREK F. ABDELZAHER
%%\affil{University of Illinois at Urbana-Champaign}}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
We consider the design of prediction market mechanisms known as automated market makers. We show that we can design these mechanisms via the mold of \emph{exponential family distributions}, a popular and well-studied probability distribution template used in statistics. We give a full development of this relationship and explore a range of benefits. We draw connections between the information aggregation of market prices and the belief aggregation of learning agents that rely on exponential family distributions. We develop a natural analysis of the market behavior as well as the price equilibrium under the assumption that the traders exhibit risk aversion according to exponential utility. We also consider similar aspects under alternative models, such as budget-constrained traders.
%
% In this paper, we draw connections between the aggregation performed by learning algorithms and the information aggregation done in prediction markets. We show that, under reasonable conditions, the behavior of rational traders can be understood as the result of performing a learning algorithm on their private data. Similarly,  the market state can be interpreted as a distribution over the outcome space. In particular, we show that a proper scoring rule can be derived from maximum entropy distributions. This scoring rule can be used as a general form of LMSR in prediction markets with over continuous outcome spaces. In order to provide insight on the behavior of rational traders in the market, we use the concept of exponential utility. We show that the traders' behavior can be understood as updating his belief using a Bayesian process and updating the market state in accordance with this utility function. These maxent prediction  markets can also be used to design  markets that are robust against adversarial traders. In fact, when traders are required to report their budgets and their beliefs, we can show that an informative trader eventually makes money and damaging traders eventually have limited influence in the market. Using ideas from convex analysis and the properties of the prediction market, we analyze the properties of the maxent market maker thus providing insight into the information content of the prediction market.
\end{abstract}

@@

%\terms{Algorithms, Economics}

\keywords{logarithmic score; exponential family; maximum entropy; risk aversion; budget constraints}

@@

Prediction markets are aggregation mechanisms that allow market prices to be interpreted as predictive probabilities on an event. Each trader in the market is assumed to have some private information that he uses to make a prediction on the outcome of the event. Traders are allowed to report their beliefs by buying and selling securities whose ultimate payoff depends on the future outcome. This will affect the state of the market, thus updating the predictive probabilities for the event. Further, since the trades are done sequentially, the trader may observe all past trades in the market and update his private beliefs based on this information. In this sense the market prices, which are in effect the prices at which the marginal trader is willing to buy or sell the available securities, can be interpreted as an aggregate ``consensus probability forecast'' of the event in question.
% For example, the Iowa Electronic Market predicts the outcome of a presidential election by allowing one to trade on securities that pay off only if the Democratic candidate wins, or only if the Republican candidate wins.

@@

In the present work we show that a number of classical statistical tools can be leveraged to design a prediction market framework in the mold of \emph{exponential family distributions}; we show that this statistical framework leads to a number of attractive properties and interpretations. Common concepts in statistics---including \emph{entropy maximization}, \emph{log loss}, and \emph{Bayesian inference}---relate to natural aspects of our class of mechanisms. In particular, the central objects in our market framework can be interpreted via concepts used to define exponential families:
\begin{itemize}
	\item the market's \emph{payoff function} corresponds to the \emph{sufficient statistics} of the distribution;
	\item the vector of \emph{outstanding shares} in the market corresponds to the \emph{natural parameter} of the distribution;
	\item the market prices correspond to \emph{mean parameters};
	\item the market's cost function corresponds to the distribution's \emph{log-partition} function.
%	\item the profit potential for a trader corresponds to a \emph{Kullback-Leibler divergence} between the trader's belief and that of the market.
\end{itemize}
% \sindhu{I don't think we have the KL divergence result here -- is this the bregman divergence in the budget section?}

We begin in Section~\ref{sec:scoring} with a discussion of scoring rules based on exponential family distributions, and we show how the framework leads to a variety of scoring rules for continuous outcome spaces.
We turn our attention to market design in Section~\ref{sec:maxentmarkets} and give a full description of our proposed mechanisms. In addition to showing the syntactic relationship between exponential families and prediction markets, we explore a number of rich semantic implications as well. In particular, we show that our formulation allows us to analyze the evolution of the market under various models of trader behavior:
\begin{itemize}
	\item Trader behavior varies depending on how they assimilate information; for example, should we consider our agents as Bayesians or frequentists. In Section~\ref{sec:bayesians} we consider traders that use a conjugate prior to update their beliefs, and we study how their trades would affect the market state.
	\item In Section~\ref{sec:exputil} we consider \emph{risk-averse} agents that optimize their bets according to exponential utility. 
	%We show how this dovetails nicely with the notion of a \emph{certainty equivalent} which is often used in risk-aversion settings. 
	In this case we can characterize precisely how a single trader interacts with the market, as well as the equilibrium reached given multiple traders; the latter result is achieved via a potential game argument. The eventual market state is a weighted combination of traders' beliefs and the initial state; the weights are proportional to risk aversion parameters.
	\item In Section~\ref{sec:budgets} we consider \emph{budget-limited traders} who are constrained in how much they influence the market. We analyze the market  under these circumstances; we are able to show that traders with good information can expect to profit and their influence over the market state increases over time whereas malicious traders have limited impact on the market.
\end{itemize}

@@

Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces. Such rules have been characterized by~\citet{Savage71}; see also~\citep{AbernethyFr12,gneiting2007strictly,lambert2008eliciting}. Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated. Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
% In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.

@@

We are interested in eliciting information about the agent's belief, in particular expectation information. Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite. The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome. A \emph{scoring rule} is a device for this purpose. Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations. Unless otherwise mentioned, all integrals including expectations are taken with respect to base measure $\nu$, which is therefore implicit in many of the formulas. A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$. The following definition is due to~\citet{lambert2008eliciting}.
%
\begin{definition} \label{def:proper-score}
A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have for all $\hmu \neq \mu$
%
\begin{equation} \label{eq-proper}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)].
\end{equation}
%
% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
\end{definition}
%

@@

It is well-known that solutions to~(\ref{maxent-prog}) are \emph{exponential family} distributions (see, for instance, \cite{WainJordan08}). These are distributions whose densities with respect to $\nu$ take the form
%
\begin{equation} \label{exp-fam}
p(x;\theta) = \exp( \la \theta, \phi(x) \ra - T(\theta) ).
\end{equation}
%
The density is stated here in terms of its \emph{natural} parametrization $\theta \in \bR^d$, where $\theta$ arises as the Lagrange multiplier associated with the linear constraints in~(\ref{maxent-prog}). The term $T(\theta)$ essentially arises as the multiplier for the normalization constraint (the density must integrate to 1), and so ensures that~(\ref{exp-fam}) is normalized:
%
\begin{equation} \label{log-part}
T(\theta) = \log \int_{\mX} \exp \la \theta, \phi(x) \ra \,d\nu(x).
\end{equation}
%
The function $T$ is known as the \emph{log-partition} or \emph{cumulant} function corresponding to the exponential family. Its domain is $\Theta = \{ \theta \in \bR^d : T(\theta) < +\infty \}$, called the natural parameter space, and we assume throughout that it is nonempty.\footnote{This may impose additional conditions on the choice of base measure $\nu$ in certain special cases. For instance, if the density over $\bR$ is known to have unit variance and statistic $\phi(x) = x$, then one uses the base measure $\nu(x) = e^{-x^2/2}$ leading to a normal distribution with fixed variance~\citep{Barndorff78}.}
%
The exponential family is \emph{regular} if $\Theta$ is open---almost all exponential families of interest, and all those we consider in this work, are regular. The family is \emph{minimal} if there is no $\alpha \in \Theta$ such that $\la \alpha, \phi(x) \ra$ is a constant over $\mX$ ($\nu$-almost everywhere); minimality is a property of the associated statistic $\phi$, usually called the \emph{sufficient statistic} in the literature.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) note
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. Note that the expected log score does not depend  on the entire maximum entropy distribution but just its expected statistic $\mu$. This is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$, where $\mu=\E_{p(x;\mu)}[\phi(x)]$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 11]{AbernethyFr12}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%
It has been understood since~\citep{mccarthy1956measures,Savage71} that scoring rules for densities and expectations can be characterized in terms of an underlying convex function. The generalized log scoring rule of Theorem~\ref{maxent-score} obtains by applying the modern characterization of~\citet{AbernethyFr12} to the optimal value function $G$ (negative maximum entropy). \citet[Thm.\ 6]{BanerjeeDhGh05} prove a similar result by showing a bijection between exponential families and Bregman divergences.

The notion of a Bregman divergence in fact provides further intuition for the result. Note that~(\ref{breg-div}) is the definition of the Bregman divergence with respect to strictly convex function $T$, written $D_T$. Therefore we have 
%
\begin{equation*}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] = D_T(\htheta,\theta) = D_G(\mu, \hmu),
\end{equation*}
%
where the last equality is a well-known identity relating the Bregman divergences of $T$ and $T^* = G$. The equation states that the agent's regret from misreporting its mean parameter does not depend on the full density $p$, only the mean $\mu$.

@@

Theorem~\ref{maxent-score} leads to a straightforward procedure for constructing score rules for expectations. Define the relevant statistic, and consider the maximum entropy (equivalently, exponential family) distribution consistent with the agent's reported mean $\mu$. The scoring rule compensates the agent according to the log likelihood of the eventual outcome according to this distribution. The interpretation is that the agent is only providing partial information about the underlying density, so the principal first infers a full density according to the principle of maximum entropy, and then scores the agent using the standard log score. We stress that the exponential families framework is a tool in the construction of scoring rules; we are not actually modeling any agent's belief via an exponential family.

An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules. By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments. The following examples illustrate the construction. Here, we have used $\mu$ and $\sigma$ for the true mean and variance and $\hmu$ and $\hsigma$ for the elicited mean and variance respectively (and $\hSigma$ for the variance of the multivariate normal).

%
\begin{example}
As a base measure we take the counting measure over the finite set $\{1,2,\ldots,d\}$, and we consider the statistic $\phi(i) \in \{0,1\}^d$ that maps outcome $i$ to the unit vector with component $i$ equal to 1. The expectation of $\phi(i)$ is simply the entire probability distribution $p$, which is a multinomial distribution (an instance of an exponential family). We recover the standard log scoring rule $S(\hat{p}, i) = \log \hat{p}(i)$. 
\end{example}
%
